---
title: "Exploration of the Exponential Distribution"
author: "Michael Szczepaniak"
date: "September 2015"
output: pdf_document
widgets: mathjax
---

### Overview  
The purpose of the report is to explore this [exponential distribution](https://en.wikipedia.org/wiki/Exponential_distribution) and verify that the distribution of samples means from this distribution conform to the [Central Limit Theorum](https://en.wikipedia.org/wiki/Illustration_of_the_central_limit_theorem).  Three items are checked as part of the verification: 1) the shape of the distribution of sample means (DoSM) approaches that of a normal distribution, 2) the mean of the DoSM approaches the mean of the original distribution, and 3) the variance of the DoSM approaches $\sigma^2/n$ where $\sigma^2$ is the variance of the original population and $n$ is the number of samples.

### Simulations
The function *createExpDoSM* created the plot shown below.  The top plot is the original exponential distribution and the bottom is the DoSM created by taking an average of 40 samples 1000 times.

```{r echo=FALSE}
library(ggplot2)
## creates simulated data:
## 1) original population from simulated exponential distribution
## 2) DoSM from 40 samples from the original population
## 3) distribution of sample standard deviations
createSimData <- function(exp.rate = 0.2) {
    data.exp.dosm <- data.frame(pop = rexp(1000, exp.rate),
                                sample.mean = rep(-1, 1000),
                                sample.sdev = rep(-1, 1000))
    for(i in 1:1000) {
        draws <- sample(data.exp.dosm$pop, 40)
        data.exp.dosm$sample.mean[i] <- mean(draws)
        data.exp.dosm$sample.sdev[i] <- sd(draws)
    }
    
    return(data.exp.dosm)
}

## creates plots of the simulated distributions
createExpDoSM <- function(mean.sim = 5, sd.sim = 5 / sqrt(40),
                          bw = 0.25, title1="", title2="") {
    library(ggplot2)
    # get frequency instead of density
    fnorm <- function(scale, ...) { return(scale * dnorm(...)) }
    
    # create some data to work with
    simData <- createSimData()
    freq.scale <- 1000 * bw
    # overlay normal frequency dist over histogram of DoSM
    p <- ggplot(simData, geom = 'blank')
    p <- p + aes(x = sample.mean, colour = 'simulation')
    # overlay normal frequency distribution
    p <- p + stat_function(fun = fnorm,
                           arg = list(scale=freq.scale,
                                      mean=mean.sim,
                                      sd=sd.sim),
                           aes(colour = 'normal'), size = 2)
    p <- p + geom_histogram(aes(x = sample.mean), alpha = 0.4, binwidth = bw)
    p <- p + scale_colour_manual(name = '', values = c('red', 'blue'))
    p <- p + theme(legend.position = c(0.85, 0.85),
                   legend.title = element_blank())
    p <- p + ggtitle(title1)
    
    print(p)
}

t1 <- "Distribution of Sample Means from Exponential Distribution"
createExpDoSM(title1 = t1)

```
Looking at the distribution of sample standard deviations...

```{r echo=FALSE}
# TODO
```


